The Strong Gravitational Lens Finding Challenge

### 2017-09-12 paper.tex
Update on Overleaf.

\title[SL Challenge]{The Strong Gravitational  Lens Finding Challenge}
%C.~E.~Petrillo$^{1}$,  C.~Tortora$^{1}$, L.~V.~E.~Koopmans$^{1}$, \and G.~Verdoes~Kleijn$^{1}$
\author[R. B. Metcalf {\it et al.}]{
R. Benton Metcalf,$^{1,4}$\thanks{E-mail: robertbenton.metcalf@unibo.it} 
M. Meneghetti,$^4$ 
Camille Avestruz,$^{11,12,13,\dagger}$
\newauthor
Fabio Bellagamba,$^1$
Cl\'ecio R. Bom,$^{18,19}$
Emmanuel Bertin,$^3$ 
\newauthor
R\'emi Cabanac,$^2$ 
Etienne Decenci\`ere,$^{10}$
R\'emi Flamary,$^8$
\newauthor
Raphael Gavazzi,$^{17}$ 
Mario Geiger,$^6$ 
Philippa Hartley,$^7$
\newauthor
Marc Huertas-Company,$^9$
Neal Jackson,$^7$ 
Eric Jullo,$^{16}$
\newauthor
Gijs Verdoes Kleijn,$^{22}$ 
Jean-Paul Kneib,$^6$
Leon Koopmans,$^{22}$
\newauthor
Fran\c{c}ois Lanusse,$^{20}$ 
Chun-Liang Li,$^{21}$
Quanbin Ma,$^{21}$ 
\newauthor
Martin Makler,$^{19}$
Nan Li,$^{14}$
Matthew Lightman,$^{15}$
\newauthor
Carlo Enrico Petrillo,$^{22}$
Christoph Sch\"{a}fer,$^6$ 
Alessandro Sonnenfeld,$^5$ 
\newauthor
Amit Tagore,$^7$
Crescenzo Tortora,$^{22}$
Diego Tuccillo,$^{9,10}$
\newauthor
Manuel B. Valent\'in,$^{19}$
and Santiago Velasco-Forero,$^{10}$
\\
% List of institutions
\vspace{0.5cm}\\
$^{1}$ Departimento di Fisica \& Astronomia, Universit\`a di Bologna, via Gabetti 93/2, 4012 Bologna, Italy \\
$^{2}$ IRAP, Universit\'e de Toulouse, CNRS, UPS, Toulouse, France.\\
$^{3}$ Sorbonne Universit\'{e}s, UPMC Univ. Paris 6 and CNRS, UMR 7095, Institut d'Astrophysique de Paris,  \\ 98bis Bd Arago, F-75014, Paris France.\\
$^{4}$ INAF-Osservatorio Astronomico di Bologna, via Ranzani 1, 40127 Bologna, Italy \\
$^{5}$ Kavli IPMU (WPI), UTIAS, The University of Tokyo, Kashiwa, Chiba 277-8583, Japan\\
$^{6}$ \red{Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL)
Observatoire de Sauverny CH - 1290 Versoix Switzerland} \\
$^{7}$ Jodrell Bank Centre for Astrophysics, School of Physics \& Astronomy, University of \\ Manchester, Oxford Rd, Manchester M13 9PL, UK \\
$^{8}$ Laboratoire Lagrange, Universi\'{e} de Nice Sophia-Antipolis, Centre National de la Recherche Scientifique,\\ Observatoire de la C\^{o}te d'Azur, Parc Valrose, 06108 Nice, France \\
$^{9}$ LERMA, Observatoire de Paris, CNRS, Universit\'e Paris Diderot, 61, Avenue de l'Observatoire F-75014, Paris, France \\
$^{10}$ MINES Paristech, PSL Research University, Centre for Mathematical Morphology, 35 rue \\ Saint-Honore, Fontainebleau, France \\
$^{11}${Enrico Fermi Institute, The University of Chicago, Chicago, IL 60637 U.S.A.}\\
$^{12}${Kavli Institute for Cosmological Physics, The University of Chicago, Chicago, IL 60637 U.S.A.}\\
$^{13}${Department of Astronomy \& Astrophysics, The University of Chicago, Chicago, IL 60637 U.S.A.}\\
$^{14}${School of Physics and Astronomy, Nottingham University, University Park, Nottingham, NG7 2RD, UK
}\\
$^{15}${JPMorgan Chase, Chicago, IL 60603 U.S.A.}\\
$^{16}$\red{ LAM - Laboratory of Astrophysics of Marseille P\^ole de l'\'Etoile Website of Ch\^ateau-Gombert 
38,} \\ \red{rue Fr\'ed\'eric Joliot-Curie 13388 Marseille cedex 13 FRANCE} \\
$^{17}$\red{Institut d'Astrophysique de Paris (IAP)- 98 bis boulevard Arago - 75014 Paris } \\
$^{18}${Centro Federal de Educa\c{c}\~ao Tecnol\'ogica Celso Suckow da Fonseca, CEP 23810-000,  Itagua\'i, RJ, Brazil}\\
$^{19}${Centro Brasileiro de Pesquisas F\'isicas, CEP 22290-180, Rio de Janeiro, RJ, Brazil}\\
$^{20}${McWilliams Center for Cosmology, Department of Physics, Carnegie Mellon University, Pittsburgh, PA 15213, USA}\\
$^{21}${School of Computer Science, Carnegie Mellon University, Pittsburgh, PA 15213, USA}\\
$^{22}${Kapteyn Astronomical Institute, University of Groningen, Postbus 800, 9700 AV, Groningen, The Netherlands}\\
$^\dagger${Provost's Postdoctoral Scholar at the University of Chicago}\\
}

@@

\subsubsection{Kapteyn Resnet (Petrillo, Tortora, Verdoes Kleijn, Koopmans) }
\red{This is a bit ragged and another submission needs to be analysed.}

Our lens-finder is based on a CNN, following the strategy adopted recently in \citep{2017arXiv170207675P}.

We decide to treat the problem as a three-class classification problem where the classes are \textit{non-lenses}, \textit{clear lenses} and \textit{dubious lenses}. We define the \textit{dubious lenses} as the lenses with lensing features with less than 160 pixels and the \textit{clear lenses} those with more than 160 pixels belonging to the lensed source. This choice is motivated by the fact that specializing the network in recognizing different class could lead to a more robust classification. In addition, in a hypothetical application of the method to real data from a survey, this could be a way to select the most blatant lenses.

Given the rules of the challenge, we had to submit a prediction for each source in the test-set as a single number: a number in the interval [0,1] or either a 0 or a 1. Unfortunately, to satisfy this requirement we had to collapse one of the class into another: we provided a 0 when the source is classified as a \textit{non-lens} and 1 if it is classified as a \textit{clear lenses} or as a \textit{dubious lenses}. Moreover, this did not allow to build a continuous ROC curve but only a binary one. \textbf{(We need to test if this was an improvement or not. I have the submission ready.)}

Ihe final submission is made using the average values of the predictions from the same network trained three different times.

@@

{\bf Architecture}
We used the CNN architecture called ResNet described in \citep{he2015deep} with three stacks of residual blocks of 5 layers.

The output layer is composed by three units. Each units gives in output a number between 0 and 1 that represents, respectively, the probability of being a non-lens, a small-lens, a big-lens.

{\bf Preprocessing}
The training fits files are preprocessed with the software \textsc{STIFF}\footnote{\href{http://www.astromatic.net/software/stiff}{\tt http://www.astromatic.net/software/stiff}} which automatically converts the fits files to gray-scale TIFF images operating a non-linear intensity transformations to enhance the low-brightness features of the image.

Due to memory limitations we down-sampled the images to 84 by 84 pixels. Most likely, this leads to a worse classification accuracy than using the original image resolution.

We augmented the training images in the following way: i) random rotation of 90, 180 or 270 degrees; ii) random shift in both $x$ and $y$ direction between -2 and +2 pixels; iii) $50\%$ probability of horizontally flipping the image.

Finally, the image border is cropped in order to have 80 by 80 pixel input images.

{\bf Training}
The network is trained by minimizing the categorical cross-entropy loss function
\begin{equation}
L = -\sum_j t_{j}\log p_{j}
\label{EQloss}
\end{equation}
where the $t_j$ and $p_j$ are respectively the label and the prediction for the class $j$. 
The minimization is done via mini-batch stochastic gradient descent with \textit{Adam} updates \citep{Kingma_2014}. We used a batch size of 56 and performed  46000 gradient updates. We started with a learning rate of $4 \times 10^{-4}$, decrease it to $4 \times 10^{-5}$ after 35000 updates and to $4 \times 10^{-6}$ after 43000 updates.
The initial values of the biases are set to zero and the weights of each filter are initialized from a random normal distribution with variance ${2/n}$ where $n$ is the number of inputs of the unit (see \citep{he2015delving}).  
We use L2-norm regularization with $\lambda= 9 \times 10^{-3}$.

>>>>

\title[SL Challenge]{The Strong Gravitational  Lens Finding Challenge}
%C.~E.~Petrillo$^{1}$,  C.~Tortora$^{1}$, L.~V.~E.~Koopmans$^{1}$, \and G.~Verdoes~Kleijn$^{1}$
\author[R. B. Metcalf {\it et al.}]{
R. Benton Metcalf,$^{1,4}$\thanks{E-mail: robertbenton.metcalf@unibo.it} 
M. Meneghetti,$^4$ 
Camille Avestruz,$^{11,12,13,\dagger}$
\newauthor
Fabio Bellagamba,$^1$
Cl\'ecio R. Bom,$^{18,19}$
Emmanuel Bertin,$^3$ 
\newauthor
R\'emi Cabanac,$^2$ 
Etienne Decenci\`ere,$^{10}$
R\'emi Flamary,$^8$
\newauthor
Raphael Gavazzi,$^{17}$ 
Mario Geiger,$^6$ 
Philippa Hartley,$^7$
\newauthor
Marc Huertas-Company,$^9$
Neal Jackson,$^7$ 
Eric Jullo,$^{16}$
\newauthor
Jean-Paul Kneib,$^6$
L\'{e}on V. E. Koopmans,$^{22}$
\newauthor
Fran\c{c}ois Lanusse,$^{20}$ 
Chun-Liang Li,$^{21}$
Quanbin Ma,$^{21}$ 
\newauthor
Martin Makler,$^{19}$
Nan Li,$^{14}$
Matthew Lightman,$^{15}$
\newauthor
Carlo Enrico Petrillo,$^{22}$
Christoph Sch\"{a}fer,$^6$ 
Alessandro Sonnenfeld,$^5$ 
\newauthor
Amit Tagore,$^7$
Crescenzo Tortora,$^{22}$
Diego Tuccillo,$^{9,10}$
\newauthor
Manuel B. Valent\'in,$^{19}$
Santiago Velasco-Forero,$^{10}$
and Gijs A. Verdoes Kleijn$^{22}$ 
\\
% List of institutions
\vspace{0.5cm}\\
$^{1}$ Departimento di Fisica \& Astronomia, Universit\`a di Bologna, via Gabetti 93/2, 4012 Bologna, Italy \\
$^{2}$ IRAP, Universit\'e de Toulouse, CNRS, UPS, Toulouse, France.\\
$^{3}$ Sorbonne Universit\'{e}s, UPMC Univ. Paris 6 and CNRS, UMR 7095, Institut d'Astrophysique de Paris,  \\ 98bis Bd Arago, F-75014, Paris France.\\
$^{4}$ INAF-Osservatorio Astronomico di Bologna, via Ranzani 1, 40127 Bologna, Italy \\
$^{5}$ Kavli IPMU (WPI), UTIAS, The University of Tokyo, Kashiwa, Chiba 277-8583, Japan\\
$^{6}$ \red{Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL)
Observatoire de Sauverny CH - 1290 Versoix Switzerland} \\
$^{7}$ Jodrell Bank Centre for Astrophysics, School of Physics \& Astronomy, University of \\ Manchester, Oxford Rd, Manchester M13 9PL, UK \\
$^{8}$ Laboratoire Lagrange, Universi\'{e} de Nice Sophia-Antipolis, Centre National de la Recherche Scientifique,\\ Observatoire de la C\^{o}te d'Azur, Parc Valrose, 06108 Nice, France \\
$^{9}$ LERMA, Observatoire de Paris, CNRS, Universit\'e Paris Diderot, 61, Avenue de l'Observatoire F-75014, Paris, France \\
$^{10}$ MINES Paristech, PSL Research University, Centre for Mathematical Morphology, 35 rue \\ Saint-Honore, Fontainebleau, France \\
$^{11}${Enrico Fermi Institute, The University of Chicago, Chicago, IL 60637 U.S.A.}\\
$^{12}${Kavli Institute for Cosmological Physics, The University of Chicago, Chicago, IL 60637 U.S.A.}\\
$^{13}${Department of Astronomy \& Astrophysics, The University of Chicago, Chicago, IL 60637 U.S.A.}\\
$^{14}${School of Physics and Astronomy, Nottingham University, University Park, Nottingham, NG7 2RD, UK
}\\
$^{15}${JPMorgan Chase, Chicago, IL 60603 U.S.A.}\\
$^{16}$\red{ LAM - Laboratory of Astrophysics of Marseille P\^ole de l'\'Etoile Website of Ch\^ateau-Gombert 
38,} \\ \red{rue Fr\'ed\'eric Joliot-Curie 13388 Marseille cedex 13 FRANCE} \\
$^{17}$\red{Institut d'Astrophysique de Paris (IAP)- 98 bis boulevard Arago - 75014 Paris } \\
$^{18}${Centro Federal de Educa\c{c}\~ao Tecnol\'ogica Celso Suckow da Fonseca, CEP 23810-000,  Itagua\'i, RJ, Brazil}\\
$^{19}${Centro Brasileiro de Pesquisas F\'isicas, CEP 22290-180, Rio de Janeiro, RJ, Brazil}\\
$^{20}${McWilliams Center for Cosmology, Department of Physics, Carnegie Mellon University, Pittsburgh, PA 15213, USA}\\
$^{21}${School of Computer Science, Carnegie Mellon University, Pittsburgh, PA 15213, USA}\\
$^{22}${Kapteyn Astronomical Institute, University of Groningen, Postbus 800, 9700 AV, Groningen, The Netherlands}\\
$^\dagger${Provost's Postdoctoral Scholar at the University of Chicago}\\
}

@@

\subsubsection{Kapteyn Resnet (Petrillo, Tortora, Verdoes Kleijn, Koopmans) }
\red{Another submission needs to be analysed.}

Our lens-finder is based on a CNN, following the strategy adopted recently in \citep{2017arXiv170207675P}. 
We decide to treat the problem as a three-class classification problem where the classes are \textit{non-lenses}, \textit{clear lenses} and \textit{dubious lenses}. We define the \textit{dubious lenses} as the lenses with lensing features with less than 160 pixels and the \textit{clear lenses} those with more than 160 pixels belonging to the lensed source. This choice is motivated by the fact that specializing the network in recognizing different class could lead to a more robust classification. In addition, in a hypothetical application of the method to real data from a survey, this could be a way to select the most blatant lenses.

Given the rules of the challenge, we had to submit a prediction for each source in the test-set as a single number: a number in the interval [0,1] or either a 0 or a 1. For satisfying this requirement we had to collapse one of the class into another: we provided a 0 when a source is classified as a \textit{non-lens} and a 1 when is classified as a \textit{clear lens} or as a \textit{dubious lenses}. This configuration did not allow to build a continuous ROC curve but only a binary one. \textbf{(We need to test if this was an improvement or not. I have the submission ready.)}
The final submission was produced averaging the values of the predictions from three CNNs with the same architecture.

@@

{\bf Architecture}
We used the CNN architecture called ResNet described in \citep{he2015deep} with three stacks of residual blocks of 5 layers. 
The output layer is composed by three units. Each units gives in output a number between 0 and 1 that represents, respectively, the probability of being a \textit{non-lens}, a \textit{dubious lens}, a \textit{certain lens}.

{\bf Preprocessing}
The training fits files are preprocessed with the software \textsc{STIFF}\footnote{\href{http://www.astromatic.net/software/stiff}{\tt http://www.astromatic.net/software/stiff}} which automatically converts the fits files to gray-scale TIFF images operating a non-linear intensity transformations to enhance the low-brightness features of the image.   
Due to memory limitations we down-sampled the images to 84 by 84 pixels. Most likely, this leads to a worse classification accuracy than using the original image resolution.
We augmented the training images in the following way: i) random rotation of 90, 180 or 270 degrees; ii) random shift in both $x$ and $y$ direction between -2 and +2 pixels; iii) $50\%$ probability of horizontally flipping the image. 
Finally, the image border is cropped in order to have 80 by 80 pixel input images.

{\bf Training}
The network is trained by minimizing the categorical cross-entropy loss function
\begin{equation}
L = -\sum_j t_{j}\log p_{j}
\label{EQloss}
\end{equation}
where the $t_j$ and $p_j$ are respectively the label and the prediction for the class $j$. 
The minimization is done via mini-batch stochastic gradient descent with \textit{Adam} updates \citep{Kingma_2014}. We used a batch size of 56 and performed  46000 gradient updates. We started with a learning rate of $4 \times 10^{-4}$, decrease it to $4 \times 10^{-5}$ after 35000 updates and to $4 \times 10^{-6}$ after 43000 updates.
The initial values of the biases are set to zero and the weights of each filter are initialized from a random normal distribution with variance ${2/n}$ where $n$ is the number of inputs of the unit \citep{He2015a}.  
We use L2-norm regularization with $\lambda= 9 \times 10^{-3}$.
